---
title: "AE 06 - Brexit - Telling Different Stories with Data"
author: "Max Asmar"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    latex_engine: xelatex
    keep_tex: false
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Introduction

In this application exercise, you'll explore how the same data can tell different stories depending on how you visualize it. You'll work with survey data about Brexit opinions across different regions of Great Britain.

**The key lesson:** The choices you make in creating visualizations (scales, proportions vs. counts, faceting vs. grouping) dramatically affect the story your visualization tells.

**In this application exercise, you will:**

- Work with factors using `fct_relevel()` and `fct_recode()`
- Create faceted visualizations
- Use free scales in faceting
- Calculate and visualize proportions
- Create grouped bar plots with `position = "dodge"`
- Format axis labels with the scales package
- Interpret how different visualizations tell different stories

**Estimated time:** 30-45 minutes

---

## Getting Started

### Fork and Clone Workflow

**Quick reminder:** Work in YOUR forked `application-exercises` repository.

1. Navigate to the `ae-06-brexit` folder in YOUR fork
2. Open `ae-06-brexit.Rmd`
3. Verify the `data` folder exists with `brexit.csv`
4. Update the YAML with your name
5. Knit to verify it works

---

## Packages
```{r load-packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(scales)  # For formatting axis labels (we'll use this in Exercise 2)
```

**What these packages do:**
- **tidyverse:** Data wrangling and visualization
- **scales:** Formatting numbers, percentages, and dates on plots

---

## The Survey

In September 2019, YouGov surveyed 1,639 GB adults with the following question:

> In hindsight, do you think Britain was right/wrong to vote to leave EU?
>
> - Right to leave  
> - Wrong to leave  
> - Don't know

The data from the survey is in `data/brexit.csv`.

---

## Load the Data
```{r load-data, message=FALSE}
brexit <- read_csv("data/brexit.csv")
```

**Explore the data:**
```{r explore-brexit}
glimpse(brexit)

# See the first few rows
head(brexit)

# Count responses by region
brexit %>%
  count(region)

# Count responses by opinion
brexit %>%
  count(opinion)
```

**Understanding the data:**

- Each row represents one survey respondent
- `region`: Which region of GB they're from
- `opinion`: Their response (Wrong, Right, Don't know)

---

## Original Visualization

In the course video, we made the following visualisation. First, we need to clean up the region names and put them in a logical order.
```{r prepare-data}
brexit <- brexit %>%
  # Reorder regions from south to north
  mutate(
    region = fct_relevel(region, "london", "rest_of_south", 
                         "midlands_wales", "north", "scot"),
    # Rename regions with proper capitalization and spacing
    region = fct_recode(region, 
                        London = "london", 
                        `Rest of South` = "rest_of_south", 
                        `Midlands / Wales` = "midlands_wales", 
                        North = "north", 
                        Scotland = "scot")
  )
```

**What this code does:**

- `fct_relevel()`: Changes the order of factor levels
- `fct_recode()`: Renames factor levels for better display
- Backticks `` `Rest of South` ``: Used for names with spaces

**Now create the original plot:**
```{r original-plot, fig.width=10, fig.height=4}
ggplot(brexit, aes(y = opinion, fill = opinion)) +
  geom_bar() +
  facet_wrap(~region, nrow = 1, labeller = label_wrap_gen(width = 12)) +
  guides(fill = "none") +
  labs(
    title = "Was Britain right/wrong to vote to leave EU?",
    subtitle = "YouGov Survey Results, 2-3 September 2019",
    caption = "Source: bit.ly/2lCJZVg",
    x = NULL, 
    y = NULL
  ) +
  scale_fill_manual(values = c(
    "Wrong" = "#ef8a62",
    "Right" = "#67a9cf",
    "Don't know" = "gray"
  )) +
  theme_minimal()
```

**What this plot shows:**

- Counts (not proportions) of each opinion
- Separated by region
- All facets use the same x-axis scale

---

## Exercise 1: Free Scales

Add `scales = "free_x"` as an argument to the `facet_wrap()` function. How does the visualisation change? How is the story this visualisation telling different than the story the original plot tells?
```{r free-scales, eval=TRUE, fig.width=10, fig.height=4}
# Remember to change eval=FALSE to eval=TRUE!

ggplot(brexit, aes(y = opinion, fill = opinion)) +
  geom_bar() +
  facet_wrap(~region,
    nrow = 1, 
    labeller = label_wrap_gen(width = 12),
    scales = "free_x"  # Add "free_x" here
  ) +
  guides(fill = "none") +
  labs(
    title = "Was Britain right/wrong to vote to leave EU?",
    subtitle = "YouGov Survey Results, 2-3 September 2019",
    caption = "Source: bit.ly/2lCJZVg",
    x = NULL, 
    y = NULL
  ) +
  scale_fill_manual(values = c(
    "Wrong" = "#ef8a62",
    "Right" = "#67a9cf",
    "Don't know" = "gray"
  )) +
  theme_minimal()
```

**What `scales = "free_x"` does:**

- Each facet gets its own x-axis scale
- The x-axis range is determined independently for each region
- Makes it easier to see proportions within each region

**Task 1.1:** How does the visualization change with free scales?

**Your answer:**

Using scales = "free_x" I am now able to have the data fill up most of each facet in the graph which allows me to better assess the distribution of "Wrong", "Right", and "Don't know" opinions for the Brexit.

**Task 1.2:** How is the story this visualization tells different from the original plot?

**Your answer:**

**Original plot emphasizes:** The original plot emphasizes the quantity of votes for each region (facet).

**Free scales plot emphasizes:** The free scales plot emphasizes the relative proportions of votes for each region (facet).

**Task 1.3:** Which visualization would be misleading if you wanted to compare the absolute number of responses across regions? Why?

**Your answer:**

The free scales plot would be misleading if I wanted to compare the absolute number of responses across regions because the free scales plot does not standardize the scale across regions. For example, in the free scales plot, at a first glance it appears that the number of votes in London was equal to the number of votes in the Rest of South whereas in the original plot the rest of the South it is significantly easier to determine that there are more votes because the bars are bigger.

---

## Exercise 2: Comparing Proportions Across Facets

First, calculate the proportion of wrong, right, and don't know answers in each category and then plot these proportions (rather than the counts). Then improve axis labeling.

**How is the story this visualisation telling different than the story the original plot tells?**

**Task 2.1:** Calculate proportions by region and opinion.
```{r calculate-proportions, eval=TRUE}
# Remember to change eval=FALSE to eval=TRUE!

brexit_prop <- brexit %>%
  count(region, opinion) %>%           # Count by region and opinion
  group_by(region) %>%             # Group by region
  mutate(prop = n / sum(n))   # Calculate proportion within each region

brexit_prop
```

**Hints:**

- Use `count(region, opinion)`
- Group by `region`
- Calculate proportion: `n / sum(n)`

**What this creates:**

- A new dataset with proportions instead of counts
- Each region's proportions sum to 1 (100%)

**Task 2.2:** Plot the proportions with improved axis labeling.
```{r plot-proportions, eval=TRUE, fig.width=10, fig.height=4}
# Remember to change eval=FALSE to eval=TRUE!

ggplot(brexit_prop, aes(y = opinion, x = prop, fill = opinion)) +
  geom_col() +                  # Use geom_col() for already-calculated values
  facet_wrap(~region, nrow = 1, labeller = label_wrap_gen(width = 12)) +
  guides(fill = "none") +
  labs(
    title = "Was Britain right/wrong to vote to leave EU?",
    subtitle = "YouGov Survey Results, 2-3 September 2019",
    caption = "Source: bit.ly/2lCJZVg",
    x = "Proportion",                  # Label for proportions
    y = NULL
  ) +
  scale_fill_manual(values = c(
    "Wrong" = "#ef8a62",
    "Right" = "#67a9cf",
    "Don't know" = "gray"
  )) +
  scale_x_continuous(labels = percent_format()) +  # Format x-axis as percentages
  theme_minimal()
```

**Hints:**

- x-axis should be `prop`
- x-axis label: something like "Proportion" or "Percentage"
- For percentage labels: `labels = percent_format()`
- Use `geom_col()` not `geom_bar()` because we already calculated the values

**Task 2.3:** How is the story this plot tells different from the original?

**Your answer:**

**Original plot (counts) shows:** The original plot shows the number of "Wrong
, "Right", and "Don't know" votes for each region.

**Proportion plot shows:** The proportion plot shows the percentage of "Wrong", "Right", and "Don't know' votes for each region, out of the total number of votes for that region.

**Which regions have the most "Wrong" responses as a proportion?** Scotland has the most ""Wrong" responses as a proportion.

---

## Exercise 3: Comparing Proportions Across Bars

Recreate the same visualisation from the previous exercise, this time dodging the bars for opinion proportions for each region, rather than faceting by region. Then improve the legend.

**How is the story this visualisation telling different than the story the previous plot tells?**

**Task 3.1:** Create a dodged bar plot.
```{r dodged-bars, eval=TRUE, fig.width=8, fig.height=5}
# Remember to change eval=FALSE to eval=TRUE!

ggplot(brexit_prop, aes(x = region, y = prop, fill = opinion)) +
  geom_col(position = "dodge") +  # "dodge" makes bars side-by-side
  labs(
    title = "Was Britain right/wrong to vote to leave EU?",
    subtitle = "YouGov Survey Results, 2-3 September 2019",
    caption = "Source: bit.ly/2lCJZVg",
    x = "Region",                  # Region label
    y = "Proportion",                  # Proportion label
    fill = "Opinion"                # Better legend title
  ) +
  scale_fill_manual(values = c(
    "Wrong" = "#ef8a62",
    "Right" = "#67a9cf",
    "Don't know" = "gray"
  )) +
  scale_y_continuous(labels = percent_format()) +  # Format y-axis as percentages
  theme_minimal()
```

**Hints:**

- x-axis: `region`
- y-axis: `prop`
- fill: `opinion`
- position: `"dodge"` (makes bars side-by-side)
- x label: "Region"
- y label: "Proportion"
- fill label: "Opinion" or "Response"
- y-axis labels: `percent_format()`

**Task 3.2:** How is this visualization different from the faceted proportion plot?

**Your answer:**

**Faceted plot makes it easier to:** The faceted plot makes it easier to compare distributions within each facet.

**Dodged plot makes it easier to:** The dodged plot makes it easier to compare distributions across regions/groups.

**Task 3.3:** Which region shows the strongest opinion one way or the other (least "Don't know")?

**Your answer:** Scotland has the smallest percentage of "Don't know" votes. Less than 10% of Scotland voters voted "Don't know".

---

## Exercise 4: Reflection on Data Storytelling

**Task 4.1:** You're writing a report arguing that Scotland is most opposed to Brexit. Which visualization would you use? Why?

**Your answer:**

I would use the dodged plot because it is very intuitive to see that Scotland has the highest proportion of "Wrong" voters. I can infer this because the highest "Wrong" bar on the dodged plot belongs to Scotland. Although the proportion plot that is not dodged from Exercise 2 also shows that Scotland has the highest proportion of "Wrong" voters, I prefer vertical graphs.

**Task 4.2:** You're writing a report comparing absolute numbers of respondents across regions. Which visualization would be most appropriate? Why?

**Your answer:**

The original plot without free scales and without dodged plot would be the most appropriate because it shows the number of voters across regions faceted by region.

**Task 4.3:** What's one thing you learned about how visualization choices affect the story you tell?

**Your answer:**

I notice that choosing to show proportions versus raw numbers can impact how readers view the disparities and comparisons across groups (regions in this case). For example, Scotland barely has any voters for this issue but it has the highest percentage of "Wrong" voters. Without knowing the number of voters, someone could misinterpret the impact of Scotland constituents on national policy.

---

## Exercise 5: Challenge (Optional)

**Task 5.1:** Create a stacked proportion bar chart where each bar represents a region and is divided by opinion, with each bar going from 0 to 100%.

**Hint:** Use `position = "fill"` in `geom_col()`.
```{r stacked-prop, eval=FALSE}
# Your code here
```

**Task 5.2:** Create a visualization showing only the proportion who said "Wrong" across regions.
```{r wrong-only, eval=FALSE}
brexit_prop %>%
  filter(opinion == "Wrong") %>%
  # Your code here
```

---

## Knit and Submit

**Before you finish:**

1. ✅ Make sure all code chunks run without errors
2. ✅ Change `eval=FALSE` to `eval=TRUE` for all exercises
3. ✅ Fill in all answer spaces
4. ✅ Knit to HTML first to see all visualizations
5. ✅ Review all plots - do they tell the story you described?

**Git workflow:**

1. Click the **Git** pane
2. Check boxes next to all changed files
3. Click **Commit**
4. Write message: "Completed AE 06 - Brexit"
5. Click **Commit** then **Push**

---

## Generate PDF for Canvas Submission

1. Click **Knit** dropdown arrow
2. Select **Knit to PDF**
3. Find `brexit.pdf` in Files pane
4. Export and save to your computer
5. Submit to Canvas

**Note:** The plots might look slightly different in PDF vs HTML due to formatting - this is normal.

---

## Troubleshooting

**Common issues:**

- **Error: "object 'brexit_prop' not found":** Make sure you ran Exercise 2 first to create this object
- **Percentage formatting not working:** Make sure you loaded the `scales` package
- **Bars not dodging:** Check that `position = "dodge"` is inside `geom_col()`
- **Wrong proportions:** Make sure you grouped by `region` before calculating proportions

**Factor issues:**
- **Regions in wrong order:** The `fct_relevel()` code sets the order
- **Region names look weird:** Check the `fct_recode()` code

**Visualization issues:**
- **Can't see all bars:** Adjust `fig.width` and `fig.height` in chunk options
- **Legend overlapping plot:** Try `theme(legend.position = "bottom")`
- **Facet labels wrapping oddly:** Adjust the `width` in `label_wrap_gen()`

**Scales package:**
- **`percent_format()` not working:** Make sure you loaded `library(scales)`
- **Percentages showing decimals:** Use `percent_format(accuracy = 1)` for whole numbers

**If you're stuck:**
- Review the walkthrough video
- Check ggplot2 documentation: `?facet_wrap`, `?geom_col`
- Try each step individually to see where errors occur
- Ask on the discussion board

---

## Key Concepts Review

**Important visualization concepts from this exercise:**

**Fixed vs. Free Scales:**
- Fixed scales: Same axis range across all facets (easier to compare absolute values)
- Free scales: Each facet has its own axis range (easier to see patterns within groups)

**Counts vs. Proportions:**
- Counts: Show absolute numbers (good for understanding sample sizes)
- Proportions: Show percentages (good for comparing across groups of different sizes)

**Faceting vs. Grouping:**
- Faceting: Separate panels for each category (easier to see patterns within categories)
- Grouping (dodged bars): All categories in one panel (easier to compare across categories)

**Data storytelling:**
- The same data can tell different stories depending on how you visualize it
- Always consider what story your visualization is telling
- Be aware of how your choices might mislead (intentionally or not)

---

## Quick Reference

**Factor manipulation:**
```r
# Reorder levels
fct_relevel(variable, "level1", "level2", ...)

# Rename levels
fct_recode(variable, 
           "New Name 1" = "old_name_1",
           "New Name 2" = "old_name_2")
```

**Calculate proportions:**
```r
data %>%
  count(group_var, count_var) %>%
  group_by(group_var) %>%
  mutate(prop = n / sum(n))
```

**Faceting:**
```r
facet_wrap(~variable, scales = "free_x")  # Free x-axis
facet_wrap(~variable, scales = "free_y")  # Free y-axis
facet_wrap(~variable, scales = "free")    # Both axes free
```

**Bar positioning:**
```r
geom_col(position = "dodge")     # Side-by-side
geom_col(position = "stack")     # Stacked (default)
geom_col(position = "fill")      # Stacked to 100%
```

**Format as percentages:**
```r
scale_x_continuous(labels = percent_format())
scale_y_continuous(labels = percent_format(accuracy = 1))  # Whole numbers
```