---
title: "AE 07 - Web Scraping IMDB"
author: "Your Name"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: show
  pdf_document:
    toc: true
    latex_engine: xelatex
    keep_tex: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, eval = FALSE)
```

## Introduction

In this application exercise, you'll learn the basics of **web scraping** - the process of extracting data from websites. You'll scrape movie and TV show data from IMDB (Internet Movie Database) to create datasets you can analyze.

**Important:** Web scraping should always be done responsibly and ethically. We're using IMDB's public charts, which are designed to be viewed. Always check a website's Terms of Service and `robots.txt` file before scraping.

**In this application exercise, you will:**

-   Read HTML pages into R
-   Extract specific elements using CSS selectors
-   Clean scraped text data with string manipulation
-   Build datasets from web-scraped data
-   Manually add data by indexing into data frames
-   Practice reproducible web scraping workflows

**Estimated time:** 60-75 minutes

**Note:** This AE uses R scripts (`.R` files) instead of R Markdown for parts of the exercise. You'll work with four scripts:

1.  `01-imdb-250movies.R` - Scrape top 250 movies (with blanks to fill)
2.  `02-imdb-tvshows.R` - Scrape popular TV shows (with blanks to fill)
3.  `03-imdb-250movies-complete.R` - Complete solution for movies
4.  `04-imdb-tvshows-complete.R` - Complete solution for TV shows

------------------------------------------------------------------------

## Getting Started

### Fork and Clone Workflow

**Quick reminder:** Work in YOUR forked `application-exercises` repository.

1.  Navigate to the `ae-07-webscraping` folder in YOUR fork
2.  You'll see this `.Rmd` file and four `.R` script files
3.  Open `ae-07-webscraping.Rmd` (this file)
4.  Update the YAML with your name
5.  Click **Knit** to verify it works

**Working with R scripts:**

R scripts (`.R` files) are different from R Markdown (`.Rmd` files):

-   No narrative text, just code and comments
-   No knitting - you run code line by line or in chunks
-   Great for data cleaning, scraping, and automation tasks
-   Comments start with `#`

------------------------------------------------------------------------

## Web Scraping Overview

### What is Web Scraping?

**Web scraping** is the process of extracting data from websites. Instead of manually copying data, we write code to:

1.  Read the HTML of a webpage
2.  Find the specific elements we want (titles, ratings, etc.)
3.  Extract the text or data from those elements
4.  Clean and organize the data into a usable format

### The `rvest` Package

We'll use the **rvest** package (pronounced "harvest") for web scraping. Key functions:

-   `read_html()`: Reads a webpage's HTML
-   `html_nodes()`: Selects specific HTML elements using CSS selectors
-   `html_text()`: Extracts text from HTML elements
-   `html_attr()`: Extracts attributes (like links) from HTML elements

### How to Find CSS Selectors

CSS selectors identify specific elements on a webpage. To find them:

1.  Open the webpage in Chrome or Firefox
2.  Right-click on the element you want → "Inspect" or "Inspect Element"
3.  In the developer tools, right-click the highlighted HTML → Copy → Copy selector
4.  Or use a browser extension like SelectorGadget

**Common CSS selector patterns:**

-   `.className` - selects elements with a specific class
-   `#idName` - selects element with a specific ID
-   `tagName` - selects all elements of that type (e.g., `a` for links)
-   Combinations like `.titleColumn a` - selects `<a>` tags inside class `titleColumn`

------------------------------------------------------------------------

## Part 1: Scraping Top 250 Movies

### Overview

We'll scrape IMDB's Top 250 Movies chart: <https://www.imdb.com/chart/top>

This page lists the highest-rated movies on IMDB. For each movie, we'll extract:

-   **Title** (e.g., "The Shawshank Redemption")
-   **Year** (e.g., 1994)
-   **Rating** (e.g., 9.3)

### Exercise 1: Inspect the Webpage

Before we code, let's understand the webpage structure.

**Task 1.1:** Open <https://www.imdb.com/chart/top> in your browser.

**Task 1.2:** Right-click on the first movie title ("The Shawshank Redemption" or whatever is #1 now) and select "Inspect" or "Inspect Element."

**Task 1.3:** Look at the HTML structure. You'll see something like:

```{html}
<td class="titleColumn">
  <a href="/title/tt0111161/">The Shawshank Redemption</a>
  <span class="secondaryInfo">(1994)</span>
</td>
```

**What patterns do you notice?**

**Your answer:**

-   Movie titles are in: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

-   Years are in: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

-   Ratings are in: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

------------------------------------------------------------------------

### Exercise 2: Open and Complete the Movies Script

**Task 2.1:** In RStudio's Files pane, open `01-imdb-250movies.R`

You'll see a script with blanks to fill in. The structure is:

1.  Load packages
2.  Read the HTML page
3.  Extract titles
4.  Extract years (with string cleaning)
5.  Extract ratings
6.  Combine into a data frame
7.  Add rank numbers

**Task 2.2:** Work through the script and fill in ALL the blanks. Hints are provided in comments.

**Key blanks to fill:**

-   `read_html("___")` - the IMDB URL
-   `html_nodes("___")` - CSS selectors for titles, years, and ratings
-   String manipulation to clean years (remove parentheses)
-   Tibble column assignments

**Task 2.3:** Run your code line by line (Ctrl+Enter or Cmd+Enter) or select all and run.

**Task 2.4:** Check your results. You should get a data frame with 250 movies. Run:

```{r check-movies, eval=FALSE}
# After running your script, these should work:
head(imdb_top_250)
nrow(imdb_top_250)  # Should be 250
```

**How many movies did you successfully scrape?** \_\_\_\_\_\_

**What is the #1 ranked movie?** \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

------------------------------------------------------------------------

### Exercise 3: Understanding the Code

Let's break down what the movies script does:

**Task 3.1:** Explain what this code does in plain English:

```{r explain-read}
page <- read_html("https://www.imdb.com/chart/top")
```

**Your answer:**

------------------------------------------------------------------------

**Task 3.2:** Explain what the pipe `%>%` does in this code:

```{r explain-pipe}
titles <- page %>%
  html_nodes(".titleColumn a") %>%
  html_text()
```

**Your answer:**

------------------------------------------------------------------------

**Task 3.3:** The year extraction uses `str_remove()` twice. Why?

```{r explain-str-remove}
years <- page %>%
  html_nodes(".secondaryInfo") %>%
  html_text() %>%
  str_remove("\\(") %>%
  str_remove("\\)") %>%
  as.numeric()
```

**Hint:** Look at the raw HTML - years appear as `(1994)`

**Your answer:**

------------------------------------------------------------------------

------------------------------------------------------------------------

### Exercise 4: Verify Your Data

**Task 4.1:** Check if your scraping worked correctly. Fill in the code:

```{r verify-movies}
# Check structure
str(imdb_top_250)

# Look at the top 10
imdb_top_250 %>%
  slice(1:10)

# Check for missing values
imdb_top_250 %>%
  summarise(
    missing_titles = sum(is.na(___)),
    missing_ratings = sum(is.na(___)),
    missing_years = sum(is.na(___))
  )
```

**How many missing values do you have?** \_\_\_\_\_\_\_

**If you have missing values, what might have gone wrong?**

------------------------------------------------------------------------

------------------------------------------------------------------------

------------------------------------------------------------------------

## Part 2: Scraping Popular TV Shows

### Overview

Now we'll scrape IMDB's Most Popular TV Shows: <https://www.imdb.com/chart/tvmeter>

This is similar to movies, but with some key differences:

-   Different CSS selectors
-   Different HTML structure
-   We'll add additional information manually for the first 3 shows

------------------------------------------------------------------------

### Exercise 5: Inspect the TV Shows Page

**Task 5.1:** Open <https://www.imdb.com/chart/tvmeter> in your browser.

**Task 5.2:** Inspect the first TV show. How is the HTML structure different from the movies page?

**Your answer:**

------------------------------------------------------------------------

------------------------------------------------------------------------

**Task 5.3:** Note the CSS selectors you'll need for:

-   TV show names: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

-   Years: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

-   Ratings: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

------------------------------------------------------------------------

### Exercise 6: Complete the TV Shows Script

**Task 6.1:** Open `02-imdb-tvshows.R` in RStudio.

**Task 6.2:** Fill in ALL the blanks in the script. This is more challenging than the movies script!

**Key differences to note:**

-   The URL is different
-   The CSS selectors are different
-   You need to extract and clean the show names (they include extra info)
-   You'll manually add genre, runtime, and episode count for the first 3 shows

**Task 6.3:** For the first 3 TV shows, visit their individual IMDB pages and record:

**Show #1:**

-   Name: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
-   Genre: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
-   Runtime (minutes): \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
-   Number of episodes: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

**Show #2:**

-   Name: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
-   Genre: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
-   Runtime (minutes): \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
-   Number of episodes: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

**Show #3:**

-   Name: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
-   Genre: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
-   Runtime (minutes): \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
-   Number of episodes: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

**Task 6.4:** Add this information to your script using indexing:

```{r manual-data-entry, eval=FALSE}
# Example for show #1
tvshows$genre[1] <- "Drama, Horror, Mystery"  # Replace with actual genre
tvshows$runtime[1] <- 494                     # Replace with actual runtime
tvshows$n_episode[1] <- 9                     # Replace with actual episodes
```

------------------------------------------------------------------------

### Exercise 7: Understanding Manual Data Entry

**Task 7.1:** Why do we use `tvshows$genre[1]` instead of creating a new data frame?

**Your answer:**

------------------------------------------------------------------------

------------------------------------------------------------------------

**Task 7.2:** What does the `[1]` mean in `tvshows$genre[1]`?

**Your answer:**

------------------------------------------------------------------------

**Task 7.3:** How would you add information for the 10th show?

**Your answer:**

------------------------------------------------------------------------

------------------------------------------------------------------------

## Part 3: Comparing Approaches

### Exercise 8: Movies vs TV Shows - What's Different?

**Task 8.1:** Compare the two scripts. What was easier/harder in the TV shows script?

**Easier in TV shows:**

------------------------------------------------------------------------

**Harder in TV shows:**

------------------------------------------------------------------------

------------------------------------------------------------------------

**Task 8.2:** Why did we need to use `str_remove_all()` and `str_squish()` for TV show names but not movie titles?

**Your answer:**

------------------------------------------------------------------------

------------------------------------------------------------------------

**Task 8.3:** Looking at the complete scripts (`03-imdb-250movies-complete.R` and `04-imdb-tvshows-complete.R`), check your work. Did you get the CSS selectors correct?

**What I got right:**

------------------------------------------------------------------------

**What I need to fix:**

------------------------------------------------------------------------

------------------------------------------------------------------------

------------------------------------------------------------------------

## Part 4: Exploring the Data

Now that you've scraped both datasets, let's analyze them!

### Exercise 9: Movie Analysis

**Task 9.1:** What decade has the most movies in the top 250?

```{r decade-analysis}
imdb_top_250 %>%
  mutate(decade = floor(year / 10) * 10) %>%
  count(decade, sort = TRUE)
```

**Answer:** The \_\_\_\_\_\_s have the most movies in the top 250.

**Task 9.2:** What's the average rating of movies in the top 250?

```{r avg-rating}
imdb_top_250 %>%
  summarise(avg_rating = mean(___))
```

**Answer:** The average rating is \_\_\_\_\_\_

**Task 9.3:** Create a visualization showing ratings over time:

```{r rating-over-time}
ggplot(imdb_top_250, aes(x = ___, y = ___)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  labs(
    title = "___",
    x = "___",
    y = "___"
  )
```

**What pattern do you see?**

------------------------------------------------------------------------

------------------------------------------------------------------------

------------------------------------------------------------------------

### Exercise 10: TV Show Analysis

**Task 10.1:** How many TV shows did you successfully scrape?

```{r count-shows}
nrow(tvshows)
```

**Answer:** \_\_\_\_\_\_ shows

**Task 10.2:** What's the range of years for the TV shows?

```{r year-range}
tvshows %>%
  summarise(
    earliest = min(___, na.rm = TRUE),
    latest = max(___, na.rm = TRUE)
  )
```

**Answer:** From \_\_\_\_\_ to \_\_\_\_\_

**Task 10.3:** Based on the 3 shows where you added genre information, what genres seem popular?

**Your answer:**

------------------------------------------------------------------------

------------------------------------------------------------------------

------------------------------------------------------------------------

## Part 5: Reflection on Web Scraping

### Exercise 11: Challenges and Ethics

**Task 11.1:** What was the most challenging part of this exercise?

**Your answer:**

------------------------------------------------------------------------

------------------------------------------------------------------------

**Task 11.2:** Why is it important to check a website's `robots.txt` file before scraping?

**Hint:** Try visiting <https://www.imdb.com/robots.txt>

**Your answer:**

------------------------------------------------------------------------

------------------------------------------------------------------------

**Task 11.3:** What are some situations where web scraping might NOT be appropriate or legal?

**Your answer:**

------------------------------------------------------------------------

------------------------------------------------------------------------

------------------------------------------------------------------------

**Task 11.4:** If IMDB changed their website design tomorrow, what would happen to your scraping code?

**Your answer:**

------------------------------------------------------------------------

------------------------------------------------------------------------

**Task 11.5:** Why might an API (Application Programming Interface) be preferable to web scraping?

**Your answer:**

------------------------------------------------------------------------

------------------------------------------------------------------------

------------------------------------------------------------------------

------------------------------------------------------------------------

## Part 6: Challenge (Optional)

### Exercise 12: Scrape Additional Information

**Challenge:** Pick one movie from the top 250 and scrape additional information from its individual page.

**Task 12.1:** Choose a movie and visit its IMDB page. What additional information is available?

**Movie chosen:** \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

**Additional info available:** (Director, Cast, Runtime, Budget, etc.)

------------------------------------------------------------------------

**Task 12.2:** Write code to scrape at least one additional piece of information:

```{r challenge-scrape}
# Your code here
```

**Task 12.3:** What CSS selector(s) did you use?

**Your answer:**

------------------------------------------------------------------------

------------------------------------------------------------------------

## Knit and Submit

**Before you finish:**

1.  ✅ Make sure both R scripts run without errors
2.  ✅ Fill in all answer spaces in this .Rmd file
3.  ✅ Knit to HTML first to verify everything works
4.  ✅ Review the HTML output

**Git workflow:**

1.  Click the **Git** pane
2.  Check boxes next to ALL changed files (this .Rmd AND the .R scripts)
3.  Click **Commit**
4.  Write a commit message: "Completed AE 07 - Web Scraping"
5.  Click **Commit**
6.  Click **Push**
7.  Verify your work is on GitHub in **YOUR fork**

------------------------------------------------------------------------

## Generate PDF for Canvas Submission

Once everything works in HTML:

1.  Click the **Knit** dropdown arrow
2.  Select **Knit to PDF**
3.  Wait for PDF generation
4.  Find `ae-07-webscraping.pdf` in the Files pane
5.  Export and save to your computer
6.  Submit the PDF to Canvas

**Remember:** Submit to BOTH GitHub (your .Rmd AND .R files) AND Canvas (the PDF)!

------------------------------------------------------------------------

## Key Concepts Summary

**Web Scraping Workflow:**

1.  **Inspect** the webpage to understand HTML structure
2.  **Identify** CSS selectors for the elements you want
3.  **Read** the HTML with `read_html()`
4.  **Extract** elements with `html_nodes()` and CSS selectors
5.  **Get text** with `html_text()`
6.  **Clean** the text with stringr functions
7.  **Combine** into a data frame
8.  **Verify** your results

**Key Functions:**

-   `read_html(url)` - Read a webpage
-   `html_nodes(css)` - Select elements by CSS selector
-   `html_text()` - Extract text from elements
-   `str_remove()`, `str_remove_all()` - Remove patterns from strings
-   `str_squish()` - Remove extra whitespace
-   `as.numeric()` - Convert text to numbers

**CSS Selectors:**

-   `.className` - Elements with a class
-   `#idName` - Element with an ID
-   `tag.className` - Specific tag with a class
-   `.class1 .class2` - Nested classes

------------------------------------------------------------------------

## Resources

**Learn more about web scraping:**

-   [rvest documentation](https://rvest.tidyverse.org/)
-   [SelectorGadget](https://selectorgadget.com/) - Browser extension for finding CSS selectors
-   [W3Schools CSS Selectors](https://www.w3schools.com/cssref/css_selectors.asp)

**Ethics and legality:**

-   Always check `robots.txt` (add `/robots.txt` to any domain)
-   Respect rate limits (don't overload servers)
-   Check Terms of Service
-   Prefer APIs when available
-   Give attribution to data sources

------------------------------------------------------------------------

## Troubleshooting

**Common issues:**

-   **Error: "could not find function":** Load the packages (tidyverse, rvest)
-   **Empty or NULL results:** CSS selector might be wrong - double-check by inspecting the page
-   **Wrong number of elements:** The selector might be matching too many or too few elements
-   **Strange text with `\n` or extra spaces:** Use `str_remove_all()` and `str_squish()` to clean
-   **Can't convert to numeric:** Make sure you removed ALL non-numeric characters first

**CSS selector issues:**

-   Use browser dev tools to test selectors
-   Start broad, then get more specific
-   Classes start with `.` and IDs start with `#`
-   Spaces in selectors mean "descendant" (nested elements)

**Website changes:**

-   Websites update their HTML frequently
-   If your code worked before but doesn't now, the selectors probably changed
-   Re-inspect the page and update your selectors

**If you're stuck:**

-   Check the complete solution scripts (`03-` and `04-`)
-   Use SelectorGadget to find CSS selectors
-   Ask on the discussion board
-   Attend office hours
-   Review the rvest documentation

------------------------------------------------------------------------

## Data Dictionary

### imdb_top_250

| Variable | Type    | Description                              |
|:---------|:--------|:-----------------------------------------|
| rank     | integer | Ranking on IMDB Top 250 list (1-250)    |
| title    | character | Movie title                             |
| rating   | numeric | IMDB rating (0-10)                       |
| year     | integer | Year of release                          |

### tvshows

| Variable  | Type    | Description                                    |
|:----------|:--------|:-----------------------------------------------|
| rank      | integer | Ranking on IMDB popularity chart (1-100)      |
| name      | character | TV show title                                |
| year      | integer | Year show premiered                            |
| score     | numeric | IMDB rating (0-10)                             |
| genre     | character | Genre(s) - manually added for first 3 shows  |
| runtime   | integer | Episode runtime in minutes                     |
| n_episode | integer | Number of episodes                             |
